### 2018-04-13 paper/appendix.tex
pass
more changes

\section{Generating a Minimal CDD Expansion}
\label{app:cons-cdd}
% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A recruions-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A minimal CDD expansion $S'$ of $S$ and
    correspondence from $S'$ to $S$.} %
  \myproc{$\expand(S)$ %
    \label{line:expand-begin} }{ %
    % auxiliary procedure:
    \myproc{$\expandaux(S')$
      \label{line:expand-aux-begin} }{ %
      % case split: optional sharing clause:
      \Switch{$\sharingclause(S')$}{ %
        % case: no siblings with overlapping descendants
        \lCase{$\none$:}{ \Return{$S'$} %
          \label{line:expand-ret} } %
        % case: siblings with 
        \lCase{$C \in S', % 
          P \in \predof{S'}$: }{ %
          % expand the return:
          \Return{$\expandaux( %
            \copyrel(S', C, P))$ } %
          \label{line:expand-recurse}
        } %
      } %
    \label{line:expand-aux-end} } %
    % base call: 
    \Return{$( \expandaux(S), \corr )$ %
      \label{line:expand-base-call}}
  } %
  %
  \caption{$\expand$:
    given a recursion-free CHC system $S$, returns a minimal
    CDD expansion $S'$ of $S$ and its correspondence.}
  \label{alg:expand}
\end{algorithm}
%
% walk through top-level procedure:
Given a recursion-free CHC system $S$, \autoref{alg:expand} returns a
minimal CDD expansion of $S$ (\autoref{defn:min-expansion}).
%
$\expand$ defines a procedure $\expandaux$
(\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
takes a CHC system $S$ and returns a minimal CDD expansion of $S$.
%
$\expand$ runs $\expandaux$ on $S$ and
returns the result, paired with the map $\corr: \predof{S'} \to \predof{S}$ 
(\autoref{line:expand-base-call}).

@@

  % contains linear systems:
  For goal \textbf{(1)}, let $S$ be an arbitrary linear
  system.
  %
  $S$ is CDD if for each clause $C$ in
  $S$, each pair of distinct predicates in
  the body of $C$, the transitive dependencies are disjoint and
  no predicates appears more than once in the body of $C$.
  (\autoref{defn:cdds}).
  %
  Let $C$ be an arbitrary clause in $S$.
  %
  Since $C$ is a linear clause, it has at most one relational
  predicate in its body. Therefore, the transitive dependencies are
  trivially disjoint and there can be no repeated predicate.
  %
  Therefore, $S$ is CDD.

@@

\section{Proof $\solvecdd$ is in co-NP}
\label{app:solve-cp}
The following is a proof of \autoref{thm:solve-cp}.
%
Namely, the upper bound of $\solvecdd$ is in co-NP.
%
\begin{proof}
  $\prectr$ and $\postctr$ construct formulas linear in the size of the
  CHC system. The satisfiability problem for the constructed formulas
  are in NP for linear arithmetic.
  %
  $\solvecdd$ at issues (at worst) a linear number of interpolation
  queries in terms of number of predicate.
  %
  Therefore, the upper bound of $\solvecdd$ is co-NP.
\end{proof}

@@

% lem: result connecting expansions to constructing solutions
For two recursion-free CHC systems $S$ and $S'$, if $S'$ is
an expansion of $S$, then the result of collapsing a
solution of $S'$ is a solution of $S$.
%
\begin{lem}
  \label{lem:expansion-sound}
  For two recursion-free CHC system $S'$ and $S$ such that $\sigma'$
  is a solution of $S'$ and $\eta$ is a correspondence from
  $S'$ to $S$, $\collapse{\eta}{\sigma'}$ is a
  solution of $S$.
\end{lem}
%
\begin{proof}
  % introduce 
  %
  For each predicate $P' \in \predof{S}'$ such that $\eta(P') = P$,
  there must exist some clause $C' \in S'$ such that $P' \in \bodyof{C'}$
  because $S'$ is an expansion of $S$.
  %
  Let predicate $Q' \in \predof{S}'$ be the head of $C'$.
  % 
  $\sigma'[P'] \land \consof{C'} \entails \sigma'[Q']$ by
  the fact that $\sigma'$ is a solution of $S'$.
  %
  Therefore,
  %
  \[ \collapse{ \eta }{ \sigma' }[P] \land \consof{C'} \entails %
  \left( \bigland_{ \substack{Q' \in \predof{S}' \\ \eta(Q') = Q} }
  \sigma(Q') = %
  \collapse{\eta}{\sigma'}[Q'] \right)
  \]
  % 
  Therefore, $\collapse{\eta}{\sigma'}$ has a solution for $P$.
  %
  Since $\collapse{\eta}{\sigma'}$ has a solution for each predicate
  in $S$, $\collapse{\eta}{\sigma'}$ is a solution of $S$.
\end{proof}

Every expansion of a solvable recursion-free CHC system is also
solvable.
% completeness of expansions:
\begin{lem}
  \label{lem:expansion-complete}
  If a recursion-free CHC system $S$ is solvable and
  $S'$ is an expansion of $S$, then $S'$ is solvable.
\end{lem}
%
\begin{proof}
  Let $\sigma$ be a solution of $S$, and let $\eta$ be a
  correspondence from $S'$ to $S$.
  %
  Let $\sigma'$ be such that for each $P' \in \predof{S}'$,
  $\sigma'(P') = \sigma(\eta(P'))$.
  %
  Then $\sigma'$ is a solution of $S'$.
\end{proof}

% CDD systems:
$\expand$ always returns a CDD expansion of its input.
%
\begin{lem}
  \label{lem:expand-corr}
  For two recursion-free CHC systems $S$ and $S'$ and a correspondence
  from $S'$ to $S$, $\eta$, such that $(S', \eta) = \expand(S)$,
  $S'$ is an CDD system and an expansion of $S$.
\end{lem}
%
\begin{proof}
  By induction over the evaluation of $\expand$ on
  an arbitrary recursion-free CHC system $S$.
  % inductive fact:
  The inductive fact is that for each evaluation step, \expandaux,
  $\corr$ is a correspondence from argument
  $S'$ to $S$.
  % base case:
  In the base case, \expandaux is called initially on $S$,
  by \autoref{alg:expand}.
  %
  $\corr$ is a correspondence from $S$ to itself, by the definition of
  $\corr$
  (\autoref{app:cons-cdd}).

@@

  % inductive case:
  In the inductive case, $\solvecdd$ calls $\solveitp$ on a predicate
  $P \in \predof{S}$ with partial solution $\sigma$.  Due to the
  topological ordering, $\sigma$ contains interpretations for each
  predicate $P \in \depsof{S}$.
  %
  Based on the definition of an interpolant(\autoref{defn:itps}),
  $\prectr(\cc{P},\sigma)$ and $\postctr(\cc{P},\sigma))$ are
  inconsistent.
  %
  The interpolant of these two formulas returned by $\solveitp$, $I$,
  is entailed by each clause $C$ where $P$ is the head where the
  predicates in $\bodyof{C}$ are substituted by their interpretations
  in $\sigma$.
  %
  $I$ is also inconsistent with all constraints that appear after $P$
  that support the query clause.
  %
  Therefore, when $\solvecdd$ updates $\sigma$ by binding $P$ to $I$
  the result is a partial solution of $S$.
\end{proof}

@@

% SolveCDD: establish completeness
\begin{lem}
  \label{lem:cdd-soln-complete}
  For a CDD system $S$ such that $S$ is
  solvable, %
  there is some $\sigma$ such that %
  $\sigma = \solvecdd(S)$.
\end{lem}
%
\begin{proof}
  % final claim:
  for all predicates $P \in \predof{S}$, $\solveitp$ returns a
  interpolant $I$, by
  \autoref{lem:vc} and the fact that $\mathcal{S}$ is solvable.
  %
  Therefore, by \autoref{lem:solve-aux} and the fact that
  $S$ is solvable, $\solvecdd(S)$ returns a
  solution of $S$.
\end{proof}

>>>>

\section{Generating a Minimal CDD Expansion}
\label{app:cons-cdd}
% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A minimal CDD expansion $S'$ of $S$ and
    a correspondence from $S'$ to $S$.} %
  \myproc{$\expand(S)$ %
    \label{line:expand-begin} }{ %
    % auxiliary procedure:
    \myproc{$\expandaux(S')$
      \label{line:expand-aux-begin} }{ %
      % case split: optional sharing clause:
      \Switch{$\sharingclause(S')$}{ %
        % case: no siblings with overlapping descendants
        \lCase{$\none$:}{ \Return{$S'$} %
          \label{line:expand-ret} } %
        % case: siblings with 
        \lCase{$C \in S', % 
          P \in \predof{S'}$: }{ %
          % expand the return:
          \Return{$\expandaux( %
            \copyrel(S', C, P))$ } %
          \label{line:expand-recurse}
        } %
      } %
    \label{line:expand-aux-end} } %
    % base call: 
    \Return{$( \expandaux(S), \corr )$ %
      \label{line:expand-base-call}}
  } %
  %
  \caption{$\expand$:
    given a recursion-free CHC system $S$, returns a minimal
    CDD expansion $S'$ of $S$ and its correspondence.}
  \label{alg:expand}
\end{algorithm}
%
% walk through top-level procedure:
Given a recursion-free CHC system $S$, \autoref{alg:expand} returns a
minimal CDD expansion of $S$ (\autoref{defn:min-expansion}).
%
$\expand$ defines a procedure $\expandaux$
(\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
takes a CHC system $S$ and returns a minimal CDD expansion of $S$.
%
$\expand$ runs $\expandaux$ on $S$ and
returns the result, paired with the map $\corr: \predof{S'} \to \predof{S}$ 
(\autoref{line:expand-base-call}).

@@

  % contains linear systems:
  For goal \textbf{(1)}, let $S$ be an arbitrary linear
  system.
  %
  $S$ is CDD if for each clause $C$ in
  $S$ \textbf{(1)} each pair of distinct predicates in
  the body of $C$ has disjoint transitive dependencies and
  \textbf{(2)} no predicate appears more than once in the body of $C$.
  (\autoref{defn:cdds}).
  %
  Let $C$ be an arbitrary clause in $S$.
  %
  Since $C$ is a linear clause, it has at most one relational
  predicate in its body. Therefore, the transitive dependencies are
  trivially disjoint and there can be no repeated predicate.
  %
  Therefore, $S$ is CDD.

@@

\section{Proof $\solvecdd$ is in co-NP}
\label{app:solve-cp}
The following is a proof of \autoref{thm:solve-cp}.
%
Namely, the upper bound of $\solvecdd$ is in co-NP.
%
\begin{proof}
  $\prectr$ and $\postctr$ construct formulas linear in the size of the
  CHC system. The satisfiability problem for the constructed formulas
  are in NP for linear arithmetic.
  %
  $\solvecdd$ issues (at worst) a linear number of interpolation
  queries in terms of number of predicate.
  %
  Therefore, the upper bound of $\solvecdd$ is co-NP.
\end{proof}

@@

% lem: result connecting expansions to constructing solutions
For two recursion-free CHC systems $S$ and $S'$, if $S'$ is
an expansion of $S$, then the result of collapsing a
solution of $S'$ is a solution of $S$.
%
\begin{lem}
  \label{lem:expansion-sound}
  For two recursion-free CHC system $S'$ and $S$ such that $\sigma'$
  is a solution of $S'$ and $\eta$ is a correspondence from
  $S'$ to $S$, $\collapse{\eta}{\sigma'}$ is a
  solution of $S$.
\end{lem}
%
\begin{proof}
  % introduce 
  %
  For each predicate $P' \in \predof{S'}$ such that $\eta(P') = P$,
  there must exist some clause $C' \in S'$ such that $P' \in \bodyof{C'}$
  because $S'$ is an expansion of $S$.
  %
  Let predicate $Q' \in \predof{S'}$ be the head of $C'$.
  % 
  $\sigma'[P'] \land \consof{C'} \entails \sigma'[Q']$ by
  the fact that $\sigma'$ is a solution of $S'$.
  %
  Therefore,
  %
  \[ \collapse{ \eta }{ \sigma' }[P] \land \consof{C'} \entails %
  \left( \bigland_{ \substack{Q' \in \predof{S'} \\ \eta(Q') = Q} }
  \sigma(Q') = %
  \collapse{\eta}{\sigma'}[Q'] \right)
  \]
  % 
  Therefore, $\collapse{\eta}{\sigma'}$ has a solution for $P$.
  %
  Since $\collapse{\eta}{\sigma'}$ has a solution for each predicate
  in $S$, $\collapse{\eta}{\sigma'}$ is a solution of $S$.
\end{proof}

Every expansion of a solvable recursion-free CHC system is also
solvable.
% completeness of expansions:
\begin{lem}
  \label{lem:expansion-complete}
  If a recursion-free CHC system $S$ is solvable and
  $S'$ is an expansion of $S$, then $S'$ is solvable.
\end{lem}
%
\begin{proof}
  Let $\sigma$ be a solution of $S$, and let $\eta$ be a
  correspondence from $S'$ to $S$.
  %
  Let $\sigma'$ be such that for each $P' \in \predof{S'}$,
  $\sigma'(P') = \sigma(\eta(P'))$.
  %
  Then $\sigma'$ is a solution of $S'$.
\end{proof}

% CDD systems:
$\expand$ always returns a CDD expansion of its input.
%
\begin{lem}
  \label{lem:expand-corr}
  For two recursion-free CHC systems $S$ and $S'$ and a correspondence
  from $S'$ to $S$, $\eta$, such that $(S', \eta) = \expand(S)$,
  $S'$ is an CDD system and an expansion of $S$.
\end{lem}
%
\begin{proof}
  By induction over the evaluation of $\expand$ on
  an arbitrary recursion-free CHC system $S$.
  % inductive fact:
  The inductive fact is that for each evaluation step
  $\corr$ is a correspondence from argument
  $S'$ to $S$.
  % base case:
  In the base case, \expandaux is called initially on $S$,
  by \autoref{alg:expand}.
  %
  $\corr$ is a correspondence from $S$ to itself, by the definition of
  $\corr$
  (\autoref{app:cons-cdd}).

@@

  % inductive case:
  In the inductive case, $\solvecdd$ calls $\solveitp$ on a predicate
  $P \in \predof{S}$ with partial solution $\sigma$.  Due to the
  topological ordering, $\sigma$ contains interpretations for each
  predicate $P \in \depsof{S}$.
  %
  Based on the definition of an interpolant (\autoref{defn:itps}),
  $\prectr(\cc{P},\sigma)$ and $\postctr(\cc{P},\sigma)$ are
  inconsistent.
  %
  The interpolant of these two formulas returned by $\solveitp$, $I$,
  is entailed by each clause $C$ where $P$ is the head where the
  predicates in $\bodyof{C}$ are substituted by their interpretations
  in $\sigma$.
  %
  $I$ is also inconsistent with all constraints that appear after $P$
  that support the query clause.
  %
  Therefore, when $\solvecdd$ updates $\sigma$ by binding $P$ to $I$
  the result is a partial solution of $S$.
\end{proof}

@@

% SolveCDD: establish completeness
\begin{lem}
  \label{lem:cdd-soln-complete}
  For a CDD system $S$ such that $S$ is
  solvable, %
  there is some $\sigma$ such that %
  $\sigma = \solvecdd(S)$.
\end{lem}
%
\begin{proof}
  % final claim:
  For all predicates $P \in \predof{S}$, $\solveitp$ returns a
  interpolant $I$, by
  \autoref{lem:vc} and the fact that $S$ is solvable.
  %
  Therefore, by \autoref{lem:solve-aux} and the fact that
  $S$ is solvable, $\solvecdd(S)$ returns a
  solution of $S$.
\end{proof}

### 2018-04-13 paper/approach.tex
pass
more changes

% walk through the approach in technical detail
\section{Technical Approach}
\label{sec:approach}
%
This section presents the technical details of our approach.
%
\autoref{sec:CDD-defn} presents the class of Clause-Dependence Disjoint
systems and its key properties.
%
\autoref{sec:solve-cdd} describes how \sys solves CDD systems
directly.
%
\autoref{sec:core-solver} describes how \sys solves a given
recursion-free system by solving an CDD system.
%
Proofs of all theorems stated in this section are in appendix
\autoref{app:char} and \autoref{app:corr}.

@@

% talk about previous system
As discussed in \autoref{sec:overview}, CDD is a superclass of the union of
class of body-disjoint systems and the class of linear systems.
%
For a given recursion-free system $S$, if each uninterpreted predicate $Q \in
\predof{S}$, appears in at most one in the body of a clause and, no
predicate appears more than once in the body of a single clause,
then $S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most one relational
predicate, then $\cc{S}$ is \emph{linear}~\cite{albarghouthi12a}.

@@

% give the general procedure:
$\shara$, given a recursion-free CHC system $S$,
constructs an CDD system $S'$.
%
\sys then directly solves $S'$ and, from this solution, constructs a solution
for $S$.
% define an expansion of a CHC system:
For two given recursion-free CHC systems $S$ and $S'$, if
there is a homomorphism from $\predof{S'}$ to $\predof{S}$ that
preserves the relationship between the clauses of $S'$ in
the clauses of $S$, then $S'$ is an expansion of
$S$ (all definitions in this section will be over fixed
$S$, and $S'$).
% formal defn:
\begin{defn}
  \label{defn:expansion}
  Let $f : \predof{S'} \to \predof{S}$ be such that
  % preserves arity:
  \textbf{(1)} for all $P' \in \predof{S'}$, $P'$ has the
  same vector of parameters as $f(P')$;
  % preserves clause structure:
  \textbf{(2)} for each clause $C' \in S'$, the clause $C$ (constructed
  by substituting all predicates $P'$ by $f(P')$) is in $S$.
  %
  Then $f$ is a \emph{correspondence} from $S'$ to $S$.
\end{defn}
%
If there is a correspondence from $S'$ to $S$, then $S'$ is an \emph{expansion}
of $S$, denoted $S \expandsto S'$.

@@

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For a recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A solution to $S$ or $\none$.}
  % expand procedure:
  \myproc{$\sys(S)$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(S', \eta) \assign %
    \expand(S)$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd(S' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma'$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
\sys (\autoref{alg:shara}), given a recursion-free CHC system $S$
(\autoref{line:shara-begin}), 
returns a solution
to $S$ or the value $\none$ to denote that $S$ is
unsolvable.
%
\sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
$S'$ of $S$.
%
\sys then invokes $\solvecdd$ on $S'$.
%
When \solvecdd~returns that $S'$ has
no solution, \sys~propagates $\none$ (\autoref{line:shara-ret-none}).
%
Otherwise, \sys constructs a solution from the CDD solution,
$\sigma'$, by invoking $\collapse{ \eta }{ \sigma' }$
(\autoref{line:shara-ret-collapse}).
%
$\textsc{Collapse}$ is designed to convert the solution for the CDD system
back to a sensible solution for the original problem. It does this by
taking the conjunction of all interpretations of predicates which
correspond to one predicate in the original problem.
%
That is, given a CDD solution $\sigma'$ and a correspondence $\eta$ from
$P' \in \predof{S'}$ to $P \in \predof{S}$,
$\collapse{\eta}{\sigma'}$ generates an entry in $\sigma$ for each
predicate in the original system:
$\sigma[P] \assign \bigland_{ \eta(P') = P} \sigma'(P')$.
%
$\expand$ is given in \autoref{app:cons-cdd}.

>>>>

% walk through the approach in technical detail
\section{Technical Approach}
\label{sec:approach}
%
This section presents the technical details of our approach.
%
\autoref{sec:CDD-defn} presents the class of Clause-Dependence Disjoint
systems and its key properties.
%
\autoref{sec:solve-cdd} describes how \sys solves CDD systems
directly.
%
\autoref{sec:core-solver} describes how \sys solves a given
recursion-free system by solving an CDD system.
%
Proofs of all theorems stated in this section are in the appendix.

@@

% talk about previous system
As discussed in \autoref{sec:overview}, CDD is a superclass of the union of
the class of body-disjoint systems and the class of linear systems.
%
For a given recursion-free system $S$, if each uninterpreted predicate $Q \in
\predof{S}$, appears in at most one in the body of a clause and, no
predicate appears more than once in the body of a single clause,
then $S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most one relational
predicate, then $\cc{S}$ is \emph{linear}~\cite{albarghouthi12a}.

@@

% give the general procedure:
$\shara$, given a recursion-free CHC system $S$,
constructs a CDD system $S'$.
%
\sys then directly solves $S'$ and, from this solution, constructs a solution
for $S$.
% define an expansion of a CHC system:
For two given recursion-free CHC systems $S$ and $S'$, if
there is a homomorphism from $\predof{S'}$ to $\predof{S}$ that
preserves the relationship between the clauses of $S'$ in
the clauses of $S$, then $S'$ is an \emph{expansion} of
$S$ (all definitions in this section will be over fixed
$S$, and $S'$).
% formal defn:
\begin{defn}
  \label{defn:expansion}
  Let $f : \predof{S'} \to \predof{S}$ be such that
  % preserves arity:
  \textbf{(1)} for all $P' \in \predof{S'}$, $P'$ has the
  same vector of parameters as $f(P')$;
  % preserves clause structure:
  \textbf{(2)} for each clause $C' \in S'$, the clause $C$ (constructed
  by substituting all predicates $P'$ by $f(P')$) is in $S$.
  %
  Then $f$ is a \emph{correspondence} from $S'$ to $S$.
\end{defn}
%
If there is a correspondence from $S'$ to $S$, then $S'$ is an \emph{expansion}
of $S$, denoted $S \expandsto S'$.

@@

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A solution to $S$ or $\none$.}
  % expand procedure:
  \myproc{$\sys(S)$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(S', \eta) \assign %
    \expand(S)$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd(S' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma'$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
\sys (\autoref{alg:shara}), given a recursion-free CHC system $S$
(\autoref{line:shara-begin}), 
returns a solution
to $S$ or the value $\none$ to denote that $S$ is
unsolvable.
%
\sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
$S'$ of $S$.
%
\sys then invokes $\solvecdd$ on $S'$.
%
When \solvecdd~returns that $S'$ has
no solution, \sys~propagates $\none$ (\autoref{line:shara-ret-none}).
%
Otherwise, \sys constructs a solution from the CDD solution,
$\sigma'$, by invoking $\collapse{ \eta }{ \sigma' }$
(\autoref{line:shara-ret-collapse}).
%
$\textsc{Collapse}$ is designed to convert the solution for the CDD system
back to a sensible solution for the original problem. It does this by
taking the conjunction of all interpretations of predicates which
correspond to one predicate in the original problem.
%
That is, given a CDD solution $\sigma'$ and a correspondence $\eta$ from
$P' \in \predof{S'}$ to $P \in \predof{S}$,
$\collapse{\eta}{\sigma'}$ generates an entry in $\sigma$ for each
predicate in the original system:
$\sigma[P] \assign \bigland_{ \eta(P') = P} \sigma'(P')$.
%
$\expand$ is given in \autoref{app:cons-cdd}.

### 2018-04-13 paper/evaluation.tex
more changes

\section{Evaluation}
\label{sec:evaluation}
% describe systems:
We performed an empirical evaluation of \sys to determine how it
compares to existing solvers for recursion-free CHC systems.
%
To do so, we implemented \sys as a modification of \duality CHC
solver, which is included in the \zthree theorem prover~\cite{z3}.
%
We modified \duality to use \sys as its solver for recursion-free CHC
systems.
%
We modified the algorithm used by \duality to generate recursion-free
unwindings of a given recursive system so that in each iteration, it
generates an unwinding with a maximal set of relational predicates and
clauses.
%
In the following context, ``\sys'' refers to \duality modified to
solve general CHC systems.

>>>>

\section{Evaluation}
\label{sec:evaluation}
% describe systems:
We performed an empirical evaluation of \sys to determine how it
compares to existing solvers for recursion-free CHC systems.
%
To do so, we implemented \sys as a modification of \duality CHC
solver, which is included in the \zthree theorem prover~\cite{z3}.
%
We modified \duality to use \sys as its solver for recursion-free CHC
systems.
%
We modified the algorithm used by \duality to generate recursion-free
unwindings of a given recursive system so that, in each iteration, it
generates an unwinding with a maximal set of relational predicates and
clauses.
%
In the following context, ``\sys'' refers to \duality modified to
solve general CHC systems.

### 2018-04-13 paper/overview.tex
more changes

% tree solver
Body-disjoint (or derivation tree)
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
a class of recursion-free CHC system where each uninterpreted
predicate appears in the body of at most one clause and appears in
such a clause exactly once.
%
Such systems cannot model a program with multiple control paths that
share a common pre-subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
body of both clause (3) and clause (4).
%
In order to solve for $\mcchc$, a solver that uses
body-disjoint systems would have to create a fresh copy of
$\cc{L}_4$. 
%
Worse, if $\cc{L}_4$ had dependencies, each would also need to be copied.

%linear systems:
Previous work has also introduced the class of linear
systems~\cite{albarghouthi12a}, where the body of each clause has at
most one uninterpreted predicate.
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls.
%
$\mcchc$ is not a linear system because the body of clause (7) has two
predicates \cc{L9} and \cc{dbl}.
%
CHC solvers that use linear systems effectively inline the
constraints for relational predicates that occur in non-linear
clauses~\cite{albarghouthi12b}.
%
In the case of $\mcchc$, inlining the constraints of $\cc{dbl}$ is
efficient, but in general such approaches can generate systems that
are exponentially larger than the input. For example, if the procedure
$\cc{dbl}$ were called more than once in $\cc{dblAbs}$ then multiple
copies of the body of the procedure would be inlined. And if the body
of this procedure were large, the inlining could become
prohibitively expensive.

>>>>

% tree solver
Body-disjoint (or derivation tree)
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
a class of recursion-free CHC system where each uninterpreted
predicate appears in the body of at most one clause and appears in
such a clause exactly once.
%
Such systems cannot model a program with multiple control paths that
share a common pre-subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
body of both clause (3) and clause (4).
%
In order to handle $\mcchc$, a solver that uses
body-disjoint systems would have to create a fresh copy of
$\cc{L}_4$. 
%
Worse, if $\cc{L}_4$ had dependencies, then each dependency would also
need to be copied.

%linear systems:
Previous work has also introduced the class of linear
systems~\cite{albarghouthi12a}, where the body of each clause has at
most one uninterpreted predicate.
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls.
%
$\mcchc$ is not a linear system because the body of clause (7) has two
predicates, \cc{L9} and \cc{dbl}.
%
CHC solvers that use linear systems effectively inline the
constraints for relational predicates that occur in non-linear
clauses~\cite{albarghouthi12b}.
%
In the case of $\mcchc$, inlining the constraints of $\cc{dbl}$ is
efficient, but in general such approaches can generate systems that
are exponentially larger than the input. For example, if the procedure
$\cc{dbl}$ were called more than once in $\cc{dblAbs}$ then multiple
copies of the body of the procedure would be inlined. And if the body
of this procedure were large, the inlining could become
prohibitively expensive.
